Multi-Task Mutual Learning for Vehicle Re-Identification

### 2019-05-08 paper.tex
Minor corrections to intro and related work sections, plus add a few more citations. Also strengthen the argument in the intro.

%%%%%%%%% ABSTRACT
\begin{abstract}
Vehicle re-identification (Re-ID) aims to search a specific vehicle
instance across non-overlapping camera views.
%
The main challenge of vehicle Re-ID is that
the visual appearance of vehicles may drastically changes
according to diverse viewpoints and illumination.
%
Most of existing vehicle Re-ID model cannot make full use of
various complementary vehicle information, e.g. vehicle type and orientation.
%
In this paper, we propose a novel {\em Multi-Task Mutual Learning} (MTML) deep model
to learn discriminative feature simultaeously from multiple branches.
%
Specially, we design a consensus learning loss function by fusing feature of the final convolutional feature maps from all branches.
%
Extensive comparative evaluations demonstrate the effectiveness of the
proposed MTML method in comparison to the state-of-the-art vehicle Re-ID techniques on the two existing benchmarks
(VehicleID and VeRi-776).
We also yield competitive performance on the NVIDIA 2019 AI City Challenge Track 2.
\end{abstract}

@@

%% Background of vehicle Re-ID
With the development of autonomous driving
and smart city applications, the need for accurately analyzing vehicles on urban streets 
by incorporating all the necessary computer vision tasks including detection, classification and
pose estimation as well as re-identification is ever-increasing.
%
Specially, vehicle re-identification attracted increasing attentions in the research community
\cite{liu2016vehicleid,liu2016veri,liu2016veri,
wang2017orientation,Wang_2017_ICCV,Zhou2018VAMI},
since it can play an important role in
intelligent transportation systems and public safety.

%% What's vehicle Re-ID?
Vehicle re-identification (Re-ID) aims at searching a specific vehicle
instance across non-overlapping camera views.
%
Due to the unavailability of license plate under uncontrolled view angles,
vehicle Re-ID by visual appearance is of great practical value in real-world applications such as smart cities.
%
This task is similar with a more popular task: person re-identification
\cite{gong2014re,chen2018person,Li2018Harmonious,xiao2016learning,
li2017person,wei2018person,song2018mask,chang2018multi,
shen2018deep,zhang2017deep,shen2018person,suh2018part},
but with more challenges:
%% Vehicle Re-ID challenges
(1) Unlike person Re-ID, pose/orientation of vehicles results in occlusion and drastic visual geometry changes, since vehicle is a kind of rigid body.
% The same vehicle under different viewpoints are visually more diverse than person,
This means that the same identity must be inferred with any given pose/orientation of a vehicle.
(2) Even in the same orientation, vehicles of different identities may look very similar
due to the same vehicle model. This requires vehicle Re-ID models
to have a more discriminative fine-grained recognition ability.

% Existing work's problem
Most of vehicle Re-ID methods \cite{liu2016vehicleid,liu2016veri,Shen_2017_ICCV,Zhou2018VAMI}
focus on using a single branch
to learn an embedding feature for vehicle instance re-identification
from the original information (e.g. the original whole vehicle image).
%
Due to the previously metioned challenges for vehicle re-identification,
this single branch structure can not take advantage of the diversity of vehicles.
%
Moreover, most existing works
\cite{liu2016vehicleid,liu2016veri}
train vehicle Re-ID deep learning model
using a single supervisory signal (e.g. vehicle ID).
%
However, vehicle ID label {\em alone} can not differentiate the differences between vehicles due to:
(1) vehicle instances of the same ID under different viewpoints,
(2) vehicle instances of different IDs under the same viewpoint.
We suggest that imposing multiple and different supervisory signals
simultaneously (e.g. vehicle ID and vehicle orientation) is more effective
for learning more fine-grained discriminative features for vehicle
Re-ID.

@@

We make two contributions in this work as follows:
(1) We formulate a novel {\em Multi-Task Mutual Learning} (MTML) deep
learning model by building each individual branch for a different
recognition task relevant to vehicle Re-ID and by taking into
consideration different four tasks including vehicle ID, multi-scale, grayscale, and orientation.
Our model aims to discover and capture concurrently complementary discriminative information.
(2) We introduce a mutual learning mechanism for improving multi-task learning robustness.
Our model benefits from multiple supervisory signals in order to
enhance model learning of more discrimative features for vehicle Re-ID.
%(3) we propose a cross-dataset joint learning framework for vehicle Re-ID.
Extensive comparative evaluations demonstrate the effectiveness of the
proposed MTML method in comparison to the state-of-the-art vehicle Re-ID techniques on the two existing benchmarks
(VehicleID \cite{liu2016vehicleid} and VeRi-776 \cite{liu2016veri}).
We also yield competitive performance on the CityFlow
\cite{tang2019cityflow} benchmark at the NVIDIA 2019 AI City Challenge.

@@

\noindent {\bf Vehicle Re-Identification. }
%Whist vehicle Re-ID is less studied than person Re-ID
%\cite{gong2014person,li2018harmonious,zhong2017camera},
%there are a handful of existing methods.
%Notably, Feris \etal \cite{feris2012large} proposed an attribute-based Re-ID method.
%The vehicles are firstly classified by different attributes like car model types and colours. The Re-ID matching is then conducted in the attribute space.
%%
%Dominik \etal \cite{zapletal2016vehicle}
%used 3D bounding boxes for rectifying car images and then concatenate colour histogram features of vehicle image pairs.
%A binary linear SVM model is then trained to verify whether a pair of images have the same identity.
%%
%Both methods rely heavily on weak hand-crafted visual features in a
%complex multi-step based approach, suffering from weak discriminative model generalisation.
Deep learning techniques have been exploited to vehicle Re-ID. For instance, Liu
\etal \cite{liu2016veri} explored a deep neural network to estimate the
visual similarities between vehicle images.
%Vehicle appearances, spatio-temporal information and license plates are
%independently and incrementally used to improve the similarity function for
%vehicle matching.  they use a deep neural network to estimate the visual
%similarities between vehicle images for enjoying the strong representation
%learning capacity.
%
Liu \etal \cite{liu2016vehicleid} also designed a Coupled Clusters Loss (CCL)
to boost a multi-branch CNN model for vehicle Re-ID.
%
All these methods utilize the global appearance features of vehicle images and
ignore local discriminative regions.
%
To explore local information motivated by the idea of landmark alignment
\cite{zhang2014facial} in both face recognition \cite{taigman2014deepface} and
human body pose estimation \cite{newell2016stacked}, Wang \etal
\cite{Wang_2017_ICCV} considered 20 vehicle keypoints for learning and aligning
local regions of a vehicle for Re-ID.  Clearly, this approach comes with extra
cost of exhaustively labeling these keypoints in a large number of vehicle
images, and the implicit assumption of having sufficient image
resolution/details for extracting these keypoints.
%

@@

\noindent{\bf Multi-Task Learning. }
Multi-task learning (MTL) is a machine learning strategy
that learns several related tasks simultaneously for
their mutual benefits \cite{argyriou2007multi}. %  \cite{evgeniou2004regularized,ando2005framework,argyriou2007multi}.
A good MTL survey with focus on neural networks
is provided in \cite{caruana1997multitask}.
%
Deep CNNs are well suited for
performing MTL as they are inherently designed to learn joint
feature representations subject to multiple label objectives
concurrently in multi-branch architectures.
%
Joint learning of multiple related tasks
has been proven to be effective in solving computer vision problems
\cite{dong2017multi,zhang2016learning}.
% \cite{dong2017multi,zhang2016learning,ahmed2008training}.
%
Critically, our method is uniquely designed to
explore the potential of MTL
in combining multiple diversities (e.g. scale and color) of the vechile image
and being supervised by multiple kinds of manual labels (e.g. ID and orientation).
Each of them is associated with an individual branch of a single model.

@@

In order to harness the benefit of all branches for the purpose of vehicle re-identification, we employ consensus learning as proposed in \cite{chen2017person} and previously harnessed for vehicle Re-ID in \cite{kanaci2018vehicle}. This is done via feature fusion of the final convolutional feature maps from all branches for consensus learning. As our branches are based on the ResNet50 architecture, these feature maps are formed via an average pooling operation which result in feature vectors of length $2048$. Hence our fused features are of length $8192$. We then add one additional fully connected layer, of size $1024$, and the output of this passed to a final identity softmax classification layer, again employed with cross entropy loss. Hence:
\begin{equation}
  p_i^C = p(\hat{y_i^C} = y_i|\mathbf{I_i}) = \frac{\exp(\hat{y_i^C})}{\sum_{k=1}^{N_{id}}\exp(\hat{y_k^C})}
\end{equation}

Additionally, we also utilise a consensus propagation mechanism, similar to the previously proposed method \cite{chen2017person,kanaci2018vehicle}. Here the consensus output is taken as 'soft targets' (as opposed to the groundtruth label 'hard targets') for the training data, and used to feedback information about the predictions made by the entire ensemble of branches. This is done concurrently with the training of the individual branches. Thie method is inspired by the idea of Knowledge Distillation (KD) \cite{hinton2015distilling}, but is different in that here we employ the combined predictions from all the `student' branches as a \emph{virtual} teacher model, rather than utilising a pre-trainined powerful teacher model to provide the soft targets.

@@

\subsection{Implementation Details}
We employ the ResNet50 \cite{} ({\bf SG: missing reference}) network
architecture as the base of our model. We train the model with
minibatches of size 8, using the Adam optimisation technique with a
learning rate of 0.0001, exponential decay rates set as follows:
$\beta_1=0.9$ and $\beta_2=0.999$. The two image sizes used were
standard $224\x224$ and small (for the scaled branch) $160\x160$.

@@

\textcolor{red}
{
Table \ref{T:benchmark_results_vehicleid} shows that ...
(Please add some comparison methods in Table 3)
}
({\bf SG: missing results})

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
Vehicle re-identification (Re-ID) aims to search a specific vehicle
instance across non-overlapping camera views.
%
The main challenge of vehicle Re-ID is that
the visual appearance of vehicles may drastically changes
according to diverse viewpoints and illumination.
%
Most of existing vehicle Re-ID model cannot make full use of
various complementary vehicle information, e.g. vehicle type and orientation.
%
In this paper, we propose a novel {\em Multi-Task Mutual Learning} (MTML) deep model
to learn discriminative features simultaeously from multiple branches.
%
Specially, we design a consensus learning loss function by fusing features from the final convolutional feature maps from all branches.
%
Extensive comparative evaluations demonstrate the effectiveness of our
proposed MTML method in comparison to the state-of-the-art vehicle Re-ID techniques on the two existing benchmarks
(VehicleID and VeRi-776).
We also yield competitive performance on the NVIDIA 2019 AI City Challenge Track 2.
\end{abstract}

@@

%% Background of vehicle Re-ID
With the development of autonomous driving
and smart city applications, the need to accurately analyze vehicles on urban streets
via multiple computer vision tasks such as detection, classification and
pose estimation, as well as re-identification, is ever-increasing.
%
Specially, vehicle re-identification has attracted increasing attention in the research community
\cite{liu2016vehicleid,liu2016veri,liu2016veri,
wang2017orientation,Wang_2017_ICCV,Zhou2018VAMI},
as it can play an important role in
intelligent transportation systems and public safety.

%% What's vehicle Re-ID?
Vehicle re-identification (Re-ID) aims to search a specific vehicle
instance across non-overlapping camera views.
%
Due to the fact that the license plate is often not visible in a number of view angles (which are generally uncontrolled),
vehicle Re-ID by visual appearance alone is of great practical value in real-world applications such as smart cities.
%
This task is similar to a more popular task: person re-identification
\cite{gong2014re,chen2018person,Li2018Harmonious,xiao2016learning,
li2017person,wei2018person,song2018mask,chang2018multi,
shen2018deep,zhang2017deep,shen2018person,suh2018part},
but with more challenges:
%% Vehicle Re-ID challenges
(1) Unlike person Re-ID, the pose/orientation of vehicles results in occlusion and drastic visual geometry changes, since the vehicle is a kind of rigid body.
% The same vehicle under different viewpoints are visually more diverse than person,
This means that is difficult to infer the same identity from any given pose/orientation of a vehicle.
(2) Even in the same orientation, vehicles of different identities may look very similar
due to the being of the same, or similar, vehicle model. This requires vehicle Re-ID models
to have a more discriminative fine-grained recognition ability.

% Existing work's problem
Most previously proposed vehicle Re-ID methods \cite{liu2016vehicleid,liu2016veri,Shen_2017_ICCV,Zhou2018VAMI} focus on using a single branch
to learn an embedded feature representation for vehicle instance re-identification
from the original information (e.g. the original whole vehicle image).
%
Due to the previously mentioned challenges for vehicle re-identification,
this single branch structure can not take advantage of the diversity of vehicles.
%
Moreover, most existing works
\cite{liu2016vehicleid,liu2016veri}
train their vehicle Re-ID deep learning model
using a single supervisory signal (e.g. vehicle ID).
%
However, we argue that vehicle ID label {\em alone} can not differentiate the differences between vehicles due to the issues we raised above. But additionally we can make use of the fact that the orientation of a vehicle alters its view in a predictable manner.
As a result, we suggest that imposing multiple and different supervisory signals
simultaneously - the vehicle ID \emph{and} vehicle orientation - allows the model to learn this variation in a more well-defined manner, and is thus more effective
for learning the fine-grained discriminative features necessary for vehicle
Re-ID. As orientation labels were provided by Wang \etal \cite{Wang_2017_ICCV} for the VeRi dataset \cite{liu2016veri}, is is possible to make use of these multiple signals for this purpose.

@@

We make two contributions in this work as follows:
(1) We formulate a novel {\em Multi-Task Mutual Learning} (MTML) deep
learning model by building each individual branch for a different
recognition task relevant to vehicle Re-ID and by taking into
consideration four different tasks: vehicle ID, multi-scale, grayscale, and orientation.
Our model aims to discover and capture concurrently the complementary discriminative information.
(2) We introduce a mutual learning mechanism for improving multi-task learning robustness.
Our model benefits from multiple supervisory signals in order to
enhance model learning of more discrimative features for vehicle Re-ID.
%(3) we propose a cross-dataset joint learning framework for vehicle Re-ID.
Extensive comparative evaluations demonstrate the effectiveness of the
proposed MTML method in comparison to the state-of-the-art vehicle Re-ID techniques on the two existing benchmarks
(VehicleID \cite{liu2016vehicleid} and VeRi-776 \cite{liu2016veri}).
We also yield competitive performance on the CityFlow
\cite{tang2019cityflow} benchmark at the NVIDIA 2019 AI City Challenge.

@@

\noindent {\bf Vehicle Re-Identification. }
%Whist vehicle Re-ID is less studied than person Re-ID
%\cite{gong2014person,li2018harmonious,zhong2017camera},
%there are a handful of existing methods.
%Notably, Feris \etal \cite{feris2012large} proposed an attribute-based Re-ID method.
%The vehicles are firstly classified by different attributes like car model types and colours. The Re-ID matching is then conducted in the attribute space.
%%
%Dominik \etal \cite{zapletal2016vehicle}
%used 3D bounding boxes for rectifying car images and then concatenate colour histogram features of vehicle image pairs.
%A binary linear SVM model is then trained to verify whether a pair of images have the same identity.
%%
%Both methods rely heavily on weak hand-crafted visual features in a
%complex multi-step based approach, suffering from weak discriminative model generalisation.
A number of deep learning techniques have been exploited for the purpose vehicle Re-ID. For instance, Liu
\etal \cite{liu2016veri} explored a deep neural network to estimate the
visual similarities between vehicle images.
%Vehicle appearances, spatio-temporal information and license plates are
%independently and incrementally used to improve the similarity function for
%vehicle matching.  they use a deep neural network to estimate the visual
%similarities between vehicle images for enjoying the strong representation
%learning capacity.
%
Liu \etal \cite{liu2016vehicleid} also designed a Coupled Clusters Loss (CCL)
to boost a multi-branch CNN model for vehicle Re-ID.
%
All these methods utilize the global appearance features of vehicle images and
ignore local discriminative regions.
%
To explore local information motivated by the idea of landmark alignment
\cite{zhang2014facial} in both face recognition \cite{taigman2014deepface} and
human body pose estimation \cite{newell2016stacked}, Wang \etal
\cite{Wang_2017_ICCV} considered 20 vehicle keypoints for learning and aligning
local regions of a vehicle for Re-ID.  Clearly, this approach comes with extra
cost of exhaustively labelling these keypoints in a large number of vehicle
images, and the implicit assumption of having sufficient image
resolution/details for extracting these keypoints.
%

@@

\noindent{\bf Multi-Task Learning. }
Multi-task learning (MTL) is a machine learning strategy
that learns several related tasks simultaneously for
their mutual benefits \cite{argyriou2007multi}. %  \cite{evgeniou2004regularized,ando2005framework,argyriou2007multi}.
A good MTL survey with focus on neural networks
is provided in \cite{caruana1997multitask}.
%
Deep CNNs are well suited for
performing MTL as they are inherently designed to learn joint
feature representations subject to multiple label objectives
concurrently in multi-branch architectures.
%
Joint learning of multiple related tasks
has been proven to be effective in solving computer vision problems
\cite{dong2017multi,zhang2016learning}.
% \cite{dong2017multi,zhang2016learning,ahmed2008training}.
%
Critically, our method is uniquely designed to
explore the potential of MTL
in combining multiple diversities (e.g. scale and color) of the vehicle image
and being supervised by multiple kinds of manual labels (e.g. ID and orientation) with each of them being associated with an individual branch of a single model.

@@

In order to harness the benefit of all branches for the purpose of vehicle re-identification, we employ consensus learning as proposed in \cite{chen2017person} and previously harnessed for vehicle Re-ID in \cite{kanaci2018vehicle}. This is done via feature fusion of the final convolutional feature maps from all branches for consensus learning. As our branches are based on the ResNet50 architecture \cite{he2016deep}, these feature maps are formed via an average pooling operation which result in feature vectors of length $2048$. Hence our fused features are of length $8192$. We then add one additional fully connected layer, of size $1024$, and the output of this passed to a final identity softmax classification layer, again employed with cross entropy loss. Hence:
\begin{equation}
  p_i^C = p(\hat{y_i^C} = y_i|\mathbf{I_i}) = \frac{\exp(\hat{y_i^C})}{\sum_{k=1}^{N_{id}}\exp(\hat{y_k^C})}
\end{equation}

Additionally, we also utilise a consensus propagation mechanism, similar to the previously proposed method \cite{chen2017person,kanaci2018vehicle}. Here the consensus output is taken as `soft targets' (as opposed to the groundtruth label `hard targets') for the training data, and used to feedback information about the predictions made by the entire ensemble of branches. This is done concurrently with the training of the individual branches. Thie method is inspired by the idea of Knowledge Distillation (KD) \cite{hinton2015distilling}, but is different in that here we employ the combined predictions from all the `student' branches as a \emph{virtual} teacher model, rather than utilising a pre-trainined powerful teacher model to provide the soft targets.

@@

\subsection{Implementation Details}
We employ the ResNet50 \cite{he2016deep} network
architecture as the base of our model. We train the model with
minibatches of size 8, using the Adam optimisation technique with a
learning rate of 0.0001, exponential decay rates set as follows:
$\beta_1=0.9$ and $\beta_2=0.999$. The two image sizes used were
standard $224\x224$ and small (for the scaled branch) $160\x160$.

@@

%% \textcolor{red}
%% {
%% Table \ref{T:benchmark_results_vehicleid} shows that ...
%% (Please add some comparison methods in Table 3)
%% }
%% ({\bf SG: missing results})
